### 2024-01-26 paper.tex
add template

\documentclass[onecolumn, 12pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{setspace}

\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{footnote}
\usepackage{endnotes}
\usepackage[bottom, multiple]{footmisc}
\usepackage[level]{fmtcount}

\usepackage{float}
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{ {./graphics} }

\usepackage[style=ieee, backend=biber, autocite=footnote]{biblatex}
\addbibresource{sources.bib}

\usepackage[pdfusetitle, hidelinks]{hyperref}

\makesavenoteenv{figure}
\SetCiteCommand{\footcite}

\title{Senior Thesis}
\author{Pedro Avalos Jim\'enez, Niklas Anderson}
\date{Wheaton College\\\today}

\begin{document}
\maketitle

\end{document}

### 2024-03-13 paper.tex
More possible sources.
add topic proposal
Work on outline

@@
 
-\usepackage[style=ieee, backend=biber, autocite=footnote]{biblatex}
+\usepackage[style=ieee, backend=biber]{biblatex}
 \addbibresource{sources.bib}
@@
 \makesavenoteenv{figure}
-\SetCiteCommand{\footcite}
+\SetCiteCommand{\cite}
 
-\title{Senior Thesis}
-\author{Pedro Avalos Jim\'enez, Niklas Anderson}
+\title{Sailing the High Seas:\\ A Christian exploration of piracy}
+\author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
 \date{Wheaton College\\\today}
@@
 
+\section{Introduction}
+
+\subsection{Background Information}
+
+\begin{displayquote}
+  \textins{T}he air in which we breathe, move, and have our physical being, incapable of
+  confinement or exclusive appropriation. Inventions then cannot, in nature, be
+  a subject of property.~\cite{barlow:wine}
+\end{displayquote}
+
+\subsubsection{Intellectual Property}
+
+\begin{enumerate}
+  \item Piracy on an individual level: Very hard to combat, because you are
+    trying to protect things that are immaterial ideas. Ideas flow freely,
+    especially in the Internet. Some people might argue that all ideas should
+    be free.
+
+  \item While some people might argue that piracy should not be allowed because
+    it stifles innovation.
+\end{enumerate}
+
+\subsection{Thesis/Argument}
+
+Piracy should not be hastily thrown aside.
+Piracy is an excellent tool for disseminating information, services, content,
+\&c. that may not have been available any other way.
+
+\section{Arguments For Piracy}
+
+\subsection{Equal Access}
+
+\begin{enumerate}
+  \item Piracy allows equal access to information
+  \item \textcquote{barlow:wine}{I am not comfortable with a model which will restrict inquiry to the wealthy.}
+
+  \item Increased prices are an artificial barrier to access certain
+    information. Piracy provides a circumvention to said artificial barrier.
+    Piracy gives consumers a way to fight systems they see as
+    unfair/censorship. Piracy is a tool of the consumer to control/balance the
+    economy~\cite{darnton:pirating-and-publishing, bohannon:everyone}
+
+  \item People pirate medical content.~\cite{till:medical-literature}
+    Medical information is valuable to everyone. Having medical information is
+    crucial to advancements in medicine, but sometimes the cost is too high.
+    The paywalls, in this case, are directly harmful to ends that are
+    beneficial to everyone.
+\end{enumerate}
+
+\subsection{Governance}
+
+\begin{enumerate}
+  \item Dissemination of information through means of piracy and torrenting
+    gives access to those within areas where Internet shutdowns are commonplace.
+
+  \item Piracy provides a way around an oppressive regime, which are on the
+    rise.~\cite{current:jigsaw}
+
+  \item IP laws arbitrarily restrict free speech / invention just because
+    someone came up with it first.~\cite[1339]{lemley:faith-based} They are
+    based on the assumption that, in doing so, they are encouraging creativity
+    (to be first to invent/create), when, in reality, creativity doesn't always
+    come first, but may rather be an improvement upon the established. The most
+    creative solution is not always the first one. IP laws and restrictions
+    block slight innovations upon the existing intellectual property.
+\end{enumerate}
+
+\subsection{Privacy}
+
+\begin{enumerate}
+  \item Piracy gives you access to cracked versions of software that provide
+    more sensible privacy/freedom than what is accessible
+    legally.~\cite{stallman:right-to-read}
+\end{enumerate}
+
+\section{Arguments Against Piracy}
+
+\subsection{Access to Dangerous/Illegal Content}
+\begin{enumerate}
+  \item Piracy generally enables free/open access to information.
+
+  \item This can be a bad thing for dangerous information that can threaten
+    lives/livelihoods.
+
+  \item Piracy would facilitate the spread of dangerous information concerning
+    national security, with few tools to mitigate and almost none to stop it.
+
+  \item Piracy provides the means to access content that is questionable/illegal.
+
+  \item When you are pirating, you need to be extremely careful with what you
+    are getting. Not all sources are trusted sources; software can be infected
+    with malware.
+
+  \item Information and content can be dispersed by bad actors.
+\end{enumerate}
+
+\subsection{Economic}
+\begin{enumerate}
+  \item Businesses are losing money because of acts of piracy.
+  \item Piracy is theft, in this sense. It is taking away from employee
+    benefits and potential tax revenue for the government.
+
+  \item Piracy takes money from creators and organizations behind content
+    creation.~\cite{congress:pirating-the-american-dream}
+
+  \item \textcquote{mrad:effects-of-ip}{According to Gould and Gruben (1996), IPRs protection stimulates economic growth if it is
+    accompanied by a policy of trade liberalization. By encouraging initiatives to innovate, IPRs
+    protection may influence the economic growth of an open country. Park and Ginarte (1997) found
+    that IPRs protection affects economic growth indirectly by stimulating the accumulation of
+  factors of production such as physical capital and R\&D capital.}
+
+    Piracy discourages innovation in the capitalist marketplace.
+\end{enumerate}
+
+\section{Christian Response}
+
+\begin{enumerate}
+  \item Drawing distinctions: Paywall to information is different from
+    withholding information for sake of security
+
+  \item Paywalls provide the means of living for artists, writers, \&c. Providing
+    ways to circumvent these when unnecessary may hinder their livelihoods.
+
+  \item Betterment of mankind as a whole, should not be behind a paywall
+
+  \item Piracy provides an avenue for missionary outreach; piracy provides
+    means of secret dissemination within an oppressive regime.
+
+  \item Piracy provides the means for access to content/information without
+    much accountability
+\end{enumerate}
+
+\section{Conclusion}
+
+\subsection{Final Thoughts / Responses}
+
+\begin{enumerate}
+  \item In response to economic argument: Would the pirate have bought the
+    product/service in the first place?
+\end{enumerate}
+
+\subsection{Applications}
+
+\subsection{Further Work}
+
+\clearpage
+\nocite{*}
+\printbibliography
 \end{document}

### 2024-03-31 paper.tex
add vim modeline and expand opening quote

@@
 
-\title{Sailing the High Seas:\\ A Christian exploration of piracy}
+\title{Sailing the High Seas:\\ A Christian Exploration of Piracy}
 \author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
@@
 \begin{displayquote}
-  \textins{T}he air in which we breathe, move, and have our physical being, incapable of
-  confinement or exclusive appropriation. Inventions then cannot, in nature, be
-  a subject of property.~\cite{barlow:wine}
+  \textins That ideas should freely spread from one to another over the globe, for the
+    moral and mutual instruction of man, and improvement of his condition, seems to have been
+    peculiarly and benevolently designed by nature, when she made them, like fire, expansible
+    over all space, without lessening their density at any point, and like the air in which we
+    breathe, move, and have our physical being, incapable of confinement or exclusive
+    appropriation. Inventions then cannot, in nature, be a subject of property.  Thomas
+    Jefferson ~\cite{barlow:wine}
 \end{displayquote}
@@
 \end{document}
+
+% vim: set ts=2 sw=2 et:

### 2024-04-01 paper.tex
add short paragraph for introduction/background

@@
 \end{displayquote}
+It is generally accepted that piracy is an illegal activity, but what is often overlooked
+is the legitimate role that piracy plays in a variety of contexts. When it comes to
+subverting oppressive regimes, for instance, the tools that piracy provides are
+indispensable and for protecting the interests of consumers from predatory business
+practices, the abilities of piracy are unmatched. Needless to say, with every unregulated
+system comes abuse and piracy is no exception. As a Christian, how should one approach
+such a tool and how can it be utilized for the betterment of God's kingdom? In the paper
+that follows, piracy will be analyzed through a number of lenses, culminating in a
+Christian response to this issue. In doing so, the hope is that the reader attains a
+fresh, wholistic understanding of the potential value of piracy.
+% explain the definition of piracy that we're using for this paper:
+% it's not just the act of downloading copyrighted content but it includes the tools that
+% are used to accomplish this task.
 

### 2024-04-08 paper.tex
add our definition of piracy to the intro section

@@
 % are used to accomplish this task.
+
+Before proceeding further into the issue of piracy, it is important to have a grasp of
+what piracy means insofar as how this paper interprets it. The dictionary defines piracy
+as ``the unauthorized use of another's production, invention, or conception especially in
+infringement of a copyright.'' While this covers the technical definition of piracy, this
+paper elects to utilize a more wholistic definition. Piracy not only includes the
+unauthorized use of someone's work, but it includes the tools and techniques that are
+commonly used to accomplish this task such as torrent trackers, VPNs, and torrent clients.
 

### 2024-04-09 paper.tex
rework and add more to intro section

@@
 indispensable and for protecting the interests of consumers from predatory business
-practices, the abilities of piracy are unmatched. Needless to say, with every unregulated
-system comes abuse and piracy is no exception. As a Christian, how should one approach
-such a tool and how can it be utilized for the betterment of God's kingdom? In the paper
-that follows, piracy will be analyzed through a number of lenses, culminating in a
-Christian response to this issue. In doing so, the hope is that the reader attains a
-fresh, wholistic understanding of the potential value of piracy.
-% explain the definition of piracy that we're using for this paper:
-% it's not just the act of downloading copyrighted content but it includes the tools that
-% are used to accomplish this task.
+practices, the abilities of piracy are unmatched.
+
+Needless to say, with every unregulated system comes abuse and piracy is no exception.
+On top of that, piracy, on an individual level, is extremely difficult to combat because
+of the immaterial nature of the items that would need to be protected. Ideas flow freely
+between people especially when the Internet is involved, so stopping this flow becomes an
+exponentially more difficult problem the more people have the restricted data. The ease
+with which ideas propagate lead people like Thomas Jefferson to claim that they are
+``incapable of confinement or exclusive appropriation.'' Consequently, two parties form:
+those that support piracy for its free and open spread of all information and those that
+see piracy as a threat to the further advancement of technology and ideas.
 
@@
 
-\subsubsection{Intellectual Property}
-
-\begin{enumerate}
-  \item Piracy on an individual level: Very hard to combat, because you are
-    trying to protect things that are immaterial ideas. Ideas flow freely,
-    especially in the Internet. Some people might argue that all ideas should
-    be free.
-
-  \item While some people might argue that piracy should not be allowed because
-    it stifles innovation.
-\end{enumerate}
+As a Christian, how should one approach such a tool and how can it be utilized for the
+betterment of God's kingdom? In the paper that follows, piracy will be analyzed through a
+number of lenses, culminating in a Christian response to this issue. In doing so, the hope
+is that the reader attains a fresh, wholistic understanding of the potential value of
+piracy.
 

### 2024-04-09 paper.tex
Start my work on writing draft.
- Convert some of the "against" section to essay/paragraph form.
- Add some notes for later
- Add some sources
- Add a quote that could be useful in our final thoughts.

--- 2024-04-08 paper.tex
@@
 \section{Arguments Against Piracy}
+While piracy enables free access to information and content, it is crucial to
+faithfully and charitably explore the potential dangers and economic
+implications associated with this practice. The arguments against piracy are
+not easy, nor should be, easy to dismiss. This paper will focus mainly on the
+ethical and economical concerns that the use of piracy introduces into society.
 
 \subsection{Access to Dangerous/Illegal Content}
-\begin{enumerate}
-  \item Piracy generally enables free/open access to information.
+Piracy has been associated with unrestricted access to information and
+content. As previously discussed, this aspect of piracy provides the means for
+those in need to access information that would otherwise be unavailable to them;
+similarly, piracy can thus provide avenues to bypass governmental restrictions
+on public information. This same benefit, however, becomes a strong argument
+against piracy when what is distributed is dangerous or illegal content.
 
-  \item This can be a bad thing for dangerous information that can threaten
-    lives/livelihoods.
+\subsubsection{Criminality of the Seeder and Leecher}
+This paper has focused on a more wholistic definition that encompasses
+the tools, techniques, motivations, and content, but the motivations and
+content that are involved in piracy should not be ignored. Piracy does involve
+a form of theft and dissemination of an individual's or organization's
+intellectual property. Piracy, thus can and most often does violate laws in
+most jurisdictions. Piracy, when it uses torrenting, involves both the
+seeder and the leecher (or the uploader and downloader, if not using torrenting).
+Depending on the legislation of the country in question, the leecher may or
+may not be found guilty of a crime, but this is complicated by the global
+context of digital piracy.
 
-  \item Piracy would facilitate the spread of dangerous information concerning
-    national security, with few tools to mitigate and almost none to stop it.
+In the United States, the downloader/streamer/leecher are generally considered
+offenders that could be subject of fines or even criminal charges. The severity
+of the penalties of piracy in these cases pose a major argument against the use
+of piracy, at least in the United States or nations where the individual
+downloading can be found guilty of a crime. ``Fair use'' and statutes of
+limitations may alleviate penalties, but repeat offences may also lose one's
+favor in these cases.~\cite{felonies.org}
 
-  \item Piracy provides the means to access content that is questionable/illegal.
+\subsubsection{Malware}
+Piracy may openly distribute dangerous illegal content, but arguably the
+larger threat is the dangerous content that remains undisclosed. Piracy
+circumvents mainline distribution channels that may incorporate secure policies
+of data integrity, authenticity, and privacy. Some piracy tools provide
+some ways to vet the authenticity of an uploader based on past uploads or based
+on the user's relationship with the piracy/torrenting
+site\footnote{%
+  \href{https://pirates-forum.org/Thread-ThePirateBay-Want-Trusted-VIP-Status-Pink-Green-Skull-See-Here}{ThePirateBay},
+  for example provides icons to display whether a user is \emph{Trusted},
+  \emph{VIP}, or \emph{Moderator}%
+}, but not all sites provide these metrics or warnings for the average end-user.
+Instead, the burden is often on the end-user, and it is not unfair to assume
+that the average person would not be able to analyze downloaded software for
+malware.
 
-  \item When you are pirating, you need to be extremely careful with what you
-    are getting. Not all sources are trusted sources; software can be infected
-    with malware.
+% TODO: Add an example of malware distributed via piracy
 
-  \item Information and content can be dispersed by bad actors.
-\end{enumerate}
+\subsubsection{Illegal content}
+Piracy, as discussed above, provides a way to distribute illegal content. This
+content is not always beneficial to society; research articles, art, movies, \&c.
+are not the only sorts of content that a pirate may distribute. Piracy opens
+the door for questionable and illegal content that may be harmful to society
+or national security.
+
+% TODO: Talk about ties to illegal organizations that are involved?
 
 \subsection{Economic}
-\begin{enumerate}
-  \item Businesses are losing money because of acts of piracy.
-  \item Piracy is theft, in this sense. It is taking away from employee
-    benefits and potential tax revenue for the government.
+Piracy, as a theft, also causes real harm to businesses. The act of piracy
+does involve stealing and distributing content from creators and organizations
+that invested time, money, and effort into the content. In this sense, piracy
+is stealing from the individuals' revenue and the government's tax
+revenue.~\cite{congress:pirating-the-american-dream}
 
-  \item Piracy takes money from creators and organizations behind content
-    creation.~\cite{congress:pirating-the-american-dream}
+\textcquote{mrad:effects-of-ip}[.]{%
+  According to Gould and Gruben (1996), IPRs protection stimulates economic
+  growth if it is accompanied by a policy of trade liberalization. By
+  encouraging initiatives to innovate, IPRs protection may influence the
+  economic growth of an open country. Park and Ginarte (1997) found that IPRs
+  protection affects economic growth indirectly by stimulating the accumulation
+  of factors of production such as physical capital and R\&D capital%
+}
+Piracy, therefore, discourages innovation in the capitalist marketplace.
 
-  \item \textcquote{mrad:effects-of-ip}{According to Gould and Gruben (1996), IPRs protection stimulates economic growth if it is
-    accompanied by a policy of trade liberalization. By encouraging initiatives to innovate, IPRs
-    protection may influence the economic growth of an open country. Park and Ginarte (1997) found
-    that IPRs protection affects economic growth indirectly by stimulating the accumulation of
-  factors of production such as physical capital and R\&D capital.}
+% TODO
+\section{Christian Response}
+A Christian response to piracy should honestly take into account both sides of
+the debate. As shown above, piracy does have some merit to it. Christians have
+to admit that piracy-promoting tools do enable
 
-    Piracy discourages innovation in the capitalist marketplace.
-\end{enumerate}
-
-\section{Christian Response}
+% 1 Peter 2:13-17
+% Romans 13:4-5
 
@@
 
+\begin{displayquote}
+  Of additional concern is the fact that by making \textins{Intellectual Property}
+  theft a national security threat without being clear about what actually
+  constitutes intellectual property, not only does the U.S. government create a
+  new reason for a militarized Internet, but it also sets the stage for companies
+  to assert that a range of other activities from file sharing to producing
+  counterfeit DVDs threaten national security and require further state
+  intervention.~\cite[p. 264]{halbert:agendas}
+\end{displayquote}
+
 \subsection{Applications}

### 2024-04-09 paper.tex
Reformat displayquore

@@
 \begin{displayquote}
-  \textins That ideas should freely spread from one to another over the globe, for the
-    moral and mutual instruction of man, and improvement of his condition, seems to have been
-    peculiarly and benevolently designed by nature, when she made them, like fire, expansible
-    over all space, without lessening their density at any point, and like the air in which we
-    breathe, move, and have our physical being, incapable of confinement or exclusive
-    appropriation. Inventions then cannot, in nature, be a subject of property.  Thomas
-    Jefferson ~\cite{barlow:wine}
+  \textins{T}hat ideas should freely spread from one to another over the globe,
+  for the moral and mutual instruction of man, and improvement of his
+  condition, seems to have been peculiarly and benevolently designed by nature,
+  when she made them, like fire, expansible over all space, without lessening
+  their density at any point, and like the air in which we breathe, move, and
+  have our physical being, incapable of confinement or exclusive appropriation.
+  Inventions then cannot, in nature, be a subject of property.
+  Thomas Jefferson~\cite{barlow:wine}
 \end{displayquote}

### 2024-04-09 paper.tex
add to open access section

@@
 \begin{displayquote}
-  \textins{T}hat ideas should freely spread from one to another over the globe,
-  for the moral and mutual instruction of man, and improvement of his
-  condition, seems to have been peculiarly and benevolently designed by nature,
-  when she made them, like fire, expansible over all space, without lessening
-  their density at any point, and like the air in which we breathe, move, and
-  have our physical being, incapable of confinement or exclusive appropriation.
-  Inventions then cannot, in nature, be a subject of property.
-  Thomas Jefferson~\cite{barlow:wine}
+  \textins That ideas should freely spread from one to another over the globe, for the
+    moral and mutual instruction of man, and improvement of his condition, seems to have been
+    peculiarly and benevolently designed by nature, when she made them, like fire, expansible
+    over all space, without lessening their density at any point, and like the air in which we
+    breathe, move, and have our physical being, incapable of confinement or exclusive
+    appropriation. Inventions then cannot, in nature, be a subject of property.  Thomas
+    Jefferson ~\cite{barlow:wine}
 \end{displayquote}
@@
 \section{Arguments For Piracy}
+As mentioned above, even though piracy is illegal, there is still merit to the tools of
+piracy and the values behind them. The merits of piracy can be divided into three realms:
+equal access to information, governance, and privacy. 
+\subsection{Equal Access}
+At its core, piracy is primarily concerned with accessing information of all kinds,
+regardless of whether someone is authorized to access that information or not. This is
+what piracy is best at and, contrary to popular belief, it is not always a bad thing.
 
-\subsection{Equal Access}
+Take the restricted access to academic journal articles, for instance. Many academic
+journals require all readers to pay not an insignificant sum to access research articles
+that could contain insight that is valuable to the human race as a whole. Because of the
+high cost of entry, certain groups are automatically excluded from reading and building
+upon the research that has already been done. John Barlow puts the situation eloquently in
+\textit{Selling Wine Without Bottles: The Economy of the Mind on the Global Net} when he
+says, \textcquote{barlow:wine}{I am not comfortable with a model which will restrict
+inquiry to the wealthy.} In the case of academic journal articles, why shouldn't anyone be
+able to better themselves by consuming quality research articles? All humans should be
+allowed to continue to learn more about God's creation without being bound by their
+financial status. Just as how someone can go to the library and checkout any book that
+tickles their fancy, they should be freely able to learn from the ongoing research of our
+world. Of course one should recognize that not all information is safe or helpful to be
+freely available on the internet, but that is a topic for a later section.
+
+A more concrete instance of where paywalls can be directly harmful to the human race is
+with regards to medical research. Till et al. stresses how important access to medical
+research is in their article \textit{Who is pirating medical literature? A bibliometric
+review of 28 million Sci-Hub downloads}. \blockcquote{till:medical-literature} {Access to
+the medical literature is essential for both the practice of evidence-based medicine and
+meaningful contribution to medical sciences. Nonetheless, only 12\% of newly published
+papers are freely accessible online, and, as of 2014, only 3 million of the 26.3 million
+articles indexed on PubMed were available on the site's repository of free materials,
+PubMed Central. Access to paywall-protected literature remains primarily through
+institutional subscriptions. Such subscriptions are costly and many struggle to afford
+access. The result is a disparity in access to the medical literature, particularly for
+those in low-income and middle-income countries (LMICs).}
+According to Till et al. the high subscription cost to access medical journal articles
+means that those in less advantaged countries cannot afford access to the material, which
+is detrimental to the furthering of medical sciences. In light of this, it should come as
+no surprise that \textcquote{till:medical-literature} {Nearly 1 million articles published
+by medical journals are downloaded on Sci-Hub each month.} Piracy enables the less
+fortunate countries to access valuable medical data that is essential to furthering
+medical research for the human race. In such a specific case as this, it should not be
+controversial to conclude that piracy is a good thing.
 
@@
 \section{Arguments Against Piracy}
-While piracy enables free access to information and content, it is crucial to
-faithfully and charitably explore the potential dangers and economic
-implications associated with this practice. The arguments against piracy are
-not easy, nor should be, easy to dismiss. This paper will focus mainly on the
-ethical and economical concerns that the use of piracy introduces into society.
 
 \subsection{Access to Dangerous/Illegal Content}
-Piracy has been associated with unrestricted access to information and
-content. As previously discussed, this aspect of piracy provides the means for
-those in need to access information that would otherwise be unavailable to them;
-similarly, piracy can thus provide avenues to bypass governmental restrictions
-on public information. This same benefit, however, becomes a strong argument
-against piracy when what is distributed is dangerous or illegal content.
+\begin{enumerate}
+  \item Piracy generally enables free/open access to information.
 
-\subsubsection{Criminality of the Seeder and Leecher}
-This paper has focused on a more wholistic definition that encompasses
-the tools, techniques, motivations, and content, but the motivations and
-content that are involved in piracy should not be ignored. Piracy does involve
-a form of theft and dissemination of an individual's or organization's
-intellectual property. Piracy, thus can and most often does violate laws in
-most jurisdictions. Piracy, when it uses torrenting, involves both the
-seeder and the leecher (or the uploader and downloader, if not using torrenting).
-Depending on the legislation of the country in question, the leecher may or
-may not be found guilty of a crime, but this is complicated by the global
-context of digital piracy.
+  \item This can be a bad thing for dangerous information that can threaten
+    lives/livelihoods.
 
-In the United States, the downloader/streamer/leecher are generally considered
-offenders that could be subject of fines or even criminal charges. The severity
-of the penalties of piracy in these cases pose a major argument against the use
-of piracy, at least in the United States or nations where the individual
-downloading can be found guilty of a crime. ``Fair use'' and statutes of
-limitations may alleviate penalties, but repeat offences may also lose one's
-favor in these cases.~\cite{felonies.org}
+  \item Piracy would facilitate the spread of dangerous information concerning
+    national security, with few tools to mitigate and almost none to stop it.
 
-\subsubsection{Malware}
-Piracy may openly distribute dangerous illegal content, but arguably the
-larger threat is the dangerous content that remains undisclosed. Piracy
-circumvents mainline distribution channels that may incorporate secure policies
-of data integrity, authenticity, and privacy. Some piracy tools provide
-some ways to vet the authenticity of an uploader based on past uploads or based
-on the user's relationship with the piracy/torrenting
-site\footnote{%
-  \href{https://pirates-forum.org/Thread-ThePirateBay-Want-Trusted-VIP-Status-Pink-Green-Skull-See-Here}{ThePirateBay},
-  for example provides icons to display whether a user is \emph{Trusted},
-  \emph{VIP}, or \emph{Moderator}%
-}, but not all sites provide these metrics or warnings for the average end-user.
-Instead, the burden is often on the end-user, and it is not unfair to assume
-that the average person would not be able to analyze downloaded software for
-malware.
+  \item Piracy provides the means to access content that is questionable/illegal.
 
-% TODO: Add an example of malware distributed via piracy
+  \item When you are pirating, you need to be extremely careful with what you
+    are getting. Not all sources are trusted sources; software can be infected
+    with malware.
 
-\subsubsection{Illegal content}
-Piracy, as discussed above, provides a way to distribute illegal content. This
-content is not always beneficial to society; research articles, art, movies, \&c.
-are not the only sorts of content that a pirate may distribute. Piracy opens
-the door for questionable and illegal content that may be harmful to society
-or national security.
-
-% TODO: Talk about ties to illegal organizations that are involved?
+  \item Information and content can be dispersed by bad actors.
+\end{enumerate}
 
 \subsection{Economic}
-Piracy, as a theft, also causes real harm to businesses. The act of piracy
-does involve stealing and distributing content from creators and organizations
-that invested time, money, and effort into the content. In this sense, piracy
-is stealing from the individuals' revenue and the government's tax
-revenue.~\cite{congress:pirating-the-american-dream}
+\begin{enumerate}
+  \item Businesses are losing money because of acts of piracy.
+  \item Piracy is theft, in this sense. It is taking away from employee
+    benefits and potential tax revenue for the government.
 
-\textcquote{mrad:effects-of-ip}[.]{%
-  According to Gould and Gruben (1996), IPRs protection stimulates economic
-  growth if it is accompanied by a policy of trade liberalization. By
-  encouraging initiatives to innovate, IPRs protection may influence the
-  economic growth of an open country. Park and Ginarte (1997) found that IPRs
-  protection affects economic growth indirectly by stimulating the accumulation
-  of factors of production such as physical capital and R\&D capital%
-}
-Piracy, therefore, discourages innovation in the capitalist marketplace.
+  \item Piracy takes money from creators and organizations behind content
+    creation.~\cite{congress:pirating-the-american-dream}
 
-% TODO
+  \item \textcquote{mrad:effects-of-ip}{According to Gould and Gruben (1996), IPRs protection stimulates economic growth if it is
+    accompanied by a policy of trade liberalization. By encouraging initiatives to innovate, IPRs
+    protection may influence the economic growth of an open country. Park and Ginarte (1997) found
+    that IPRs protection affects economic growth indirectly by stimulating the accumulation of
+  factors of production such as physical capital and R\&D capital.}
+
+    Piracy discourages innovation in the capitalist marketplace.
+\end{enumerate}
+
 \section{Christian Response}
-A Christian response to piracy should honestly take into account both sides of
-the debate. As shown above, piracy does have some merit to it. Christians have
-to admit that piracy-promoting tools do enable
-
-% 1 Peter 2:13-17
-% Romans 13:4-5
 
@@
 
-\begin{displayquote}
-  Of additional concern is the fact that by making \textins{Intellectual Property}
-  theft a national security threat without being clear about what actually
-  constitutes intellectual property, not only does the U.S. government create a
-  new reason for a militarized Internet, but it also sets the stage for companies
-  to assert that a range of other activities from file sharing to producing
-  counterfeit DVDs threaten national security and require further state
-  intervention.~\cite[p. 264]{halbert:agendas}
-\end{displayquote}
-
 \subsection{Applications}

### 2024-04-09 paper.tex
finish off open access section

@@
 controversial to conclude that piracy is a good thing.
+
+Medical journals are not the only source of information that can sit behind a paywall.
+Sometimes, as in the previous case, there is information that one might argue is unjustly
+kept away from the common man. What one considers to be unjust is subjective, but the idea
+is that piracy is a tool that consumers can employ to fight systems that they deem to be
+unjust. Without piracy, one might have to roll over and accept their limited access or
+resort to a more extreme and likely illegal approach. To put it another way, piracy acts
+as a check on businesses to discourage them from integrating unpopular, anti-consumer
+practices into their business model. When a consumer wants to purchase something and they
+cannot afford it, they might start to look for alternative way of obtaining it. But, when
+they can afford it and they think it is fairly priced, they will purchase it. Thus, piracy
+is often where tech-savvy persons turn to when they need something that they feel is
+exorbitantly priced, but is not often utilized when they feel the purchase is worthwhile.
+This way, piracy acts as a safeguard for consumers when they feel they are being taken
+advantage of by predatory business practices.
+% XXX the statements in the paragraph above are largely unsubstantiated >:D
+% XXX add darnton:pirating-and-publishing and bohannon:everyone sources if needed
 

### 2024-04-09 paper.tex
add additional paragraphs in governance

@@
 
-\begin{enumerate}
-  \item Piracy allows equal access to information
-  \item \textcquote{barlow:wine}{I am not comfortable with a model which will restrict inquiry to the wealthy.}
+\subsection{Governance}
+Given that piracy is an illegal usage of another person's conception, it has much to say
+in the way of governance and intellectual property laws. The tools of piracy provide the
+means of subverting an oppressive regime and call into question the legitimacy of
+intellectual property laws.
 
-  \item Increased prices are an artificial barrier to access certain
-    information. Piracy provides a circumvention to said artificial barrier.
-    Piracy gives consumers a way to fight systems they see as
-    unfair/censorship. Piracy is a tool of the consumer to control/balance the
-    economy~\cite{darnton:pirating-and-publishing, bohannon:everyone}
+Having a very firm grip on what citizens can see on the internet is one of the key
+characteristics of a despotic government. Such governments know that if they can control
+the flow of information on the Internet, it becomes much easier to keep the populace under
+control. The government's imposing of restrictions on the Internet is clearly seen when
+\textcquote{current:jigsaw} {Tunisia significantly ramped up its already aggressive
+blocking of specific websites in response to unrest that would ultimately unseat its
+government.} Even more extreme is when governments shutdown the internet entirely to
+quickly cease the spread of information. This is exactly what happened on January 28, 2011
+in Egypt: \textcquote{current:jigsaw} {Almost simultaneously, about 3,500 individual
+Border Gateway Protocol routes [...] were withdrawn on orders from the Egyptian
+government, cutting the country off from the rest of the world and bringing internal
+communication to a halt.} Fortunately, as already mentioned earlier, piracy excels at
+spreading information even when there are those that do not want it spread. BitTorrent is
+a peer-to-peer file sharing protocol that is often used when pirating content on the
+Internet. A peer-to-peer protocol, by its very nature, does not require a centralized
+server to store all the files that are to be shared. Instead, the source of the file is
+each person who has the file downloaded. What this means for those under the rule of an
+oppressive government is that they can spread information to others in such a way that the
+government cannot easily block its source. Pieces of the information are sent to the one
+downloading the file from everyone who had the file. There is no single source of the
+information that the government can block to stop the spread. By a similar token, there
+are peer-to-peer instant messaging applications that do not require an internet connection
+to function; you need only be in the vicinity of other people that the message can hop
+through to get to the destination. Both of these tools, in the context of piracy, facilitate
+the illegal spread of information, but in the context of a despotic government, can be
+employed to subvert their attempts at quelling the dispersal of information that could be
+harmful to their rule.
 
-  \item People pirate medical content.~\cite{till:medical-literature}
-    Medical information is valuable to everyone. Having medical information is
-    crucial to advancements in medicine, but sometimes the cost is too high.
-    The paywalls, in this case, are directly harmful to ends that are
-    beneficial to everyone.
-\end{enumerate}
-
-\subsection{Governance}
+Regarding piracy and intellectual property laws, it is clear that the two are at odds with
+one another. That being said, it is worth examining whether IP laws accomplish what they
+seek to do: encourage innovation. Mark Lemley presents an argument against IP laws when he
+writes, \blockcquote[1339]{lemley:faith-based}{[IP] intervenes in the market to interfere
+with the freedom of others to do what they want in hopes of achieving the end of
+encouraging creativity. If we take that purpose out of the equation, we are left with a
+belief system that says the government should restrict your speech and freedom of action
+in favor of mine, not because doing so will improve the world, but simply because I spoke
+first.} 
+To put it differently, intellectual property laws are based on the assumption that in
+restricting who is allowed to have certain information, they are encouraging creativity
+when, in reality, creativity does not always come first. IP laws are supposed to make
+people want to work hard to get a good idea that they can have exclusive ownership of, but
+having exclusive rights to an idea stifles further innovation. Under this system, only the
+owner has the right to make changes. In fact, it seems that there would be more creativity
+without IP laws. If everything was open to scrutiny and replication, there is bound to be
+some improvement eventually because of all the people that could have a hand in making it
+better. If a big streaming platform were to forfeit one of their IPs to the public, there
+is no telling what kind of content could be created. In this way, piracy could allow for
+the incremental improvement and scrutiny of conceptions, thus making them better.
 

### 2024-04-10 paper.tex
add privacy subsection

@@
 
-\begin{enumerate}
-  \item Dissemination of information through means of piracy and torrenting
-    gives access to those within areas where Internet shutdowns are commonplace.
-
-  \item Piracy provides a way around an oppressive regime, which are on the
-    rise.~\cite{current:jigsaw}
-
-  \item IP laws arbitrarily restrict free speech / invention just because
-    someone came up with it first.~\cite[1339]{lemley:faith-based} They are
-    based on the assumption that, in doing so, they are encouraging creativity
-    (to be first to invent/create), when, in reality, creativity doesn't always
-    come first, but may rather be an improvement upon the established. The most
-    creative solution is not always the first one. IP laws and restrictions
-    block slight innovations upon the existing intellectual property.
-\end{enumerate}
-
 \subsection{Privacy}
-
-\begin{enumerate}
-  \item Piracy gives you access to cracked versions of software that provide
-    more sensible privacy/freedom than what is accessible
-    legally.~\cite{stallman:right-to-read}
-\end{enumerate}
+As a side note, there are instances where pirated software can be more privacy-oriented
+than the paid version. Sometimes the pirated version is ``cracked'' to remove certain
+digital rights management (DRM) tools that often run background tasks unbeknownst to the
+user. These background tasks often leak bits of information about the user in order to
+validate that they are legitimate owners of the software. By collecting information about
+users, a business invades the privacy of its user base. However, pirated and cracked
+software does not have such DRM because it needed to be removed to allow it to be freely
+distributed. Thus, when it comes to software that contains aggressive DRM, piracy offers
+better security by default. Unfortunately, some pirated software introduces malicious
+code, a subject that will be explored in the next section.
 

### 2024-04-10 paper.tex
Work on Against and Christian Response sections.
Still need to add more to the Christian response (and include it more
throughout the essay). I also still need to add more to the Against
section, not just the malware section.

@@
 
-\begin{displayquote}
+\blockcquote{barlow:wine}{%
   \textins{T}hat ideas should freely spread from one to another over the globe,
@@
   Inventions then cannot, in nature, be a subject of property.
-  Thomas Jefferson~\cite{barlow:wine}
-\end{displayquote}
+  \emph{Thomas Jefferson}%
+}
+
 It is generally accepted that piracy is an illegal activity, but what is often overlooked
@@
 in Egypt: \textcquote{current:jigsaw} {Almost simultaneously, about 3,500 individual
-Border Gateway Protocol routes [...] were withdrawn on orders from the Egyptian
+Border Gateway Protocol routes \textelp{} were withdrawn on orders from the Egyptian
 government, cutting the country off from the rest of the world and bringing internal
@@
 \subsubsection{Malware}
+The intersection of trust and computing is not a new topic. Ken Thompson, the
+creator of the Unix operating system, made this the topic of his Turing Award
+lecture. In his lecture, he poses the question,
+\textcquote[761]{thompson:trust}[.]{%
+  to what extent should one trust a statement that
+  a program is free of Trojan Horses? Perhaps it is more important to trust the
+  people who wrote the software%
+} With piracy, however, we add an additional actor in this chain of trust.
+Thompson's lecture focuses solely on the developer-to-user relationship of
+trust, with the assumption of an open-source project. Thompson, thus, comments
+on how \textcquote[763]{thompson:trust}{%
+  \textins{s}uch blatant code would not go undetected for long. Even the most
+  casual perusal of the source of the C compiler would raise suspicions%
+} When piracy is the method of distribution, however, the assumption of
+equality between the source code and the delivered software does not hold. The
+pirated software may have been tampered with, and the user has no simple way to
+detect changes in the file(s) without doing some (often) complicated forensic
+work. Although anti-malware tools may mitigate these issues, the problem of
+trust still holds, and the anti-malware tools are not perfect. The character of
+the seeder must be questioned, and it already stands on shaky ground if the
+legality of their actions speaks for their moral compass and trustworthiness.
+
 Piracy may openly distribute dangerous illegal content, but arguably the
@@
 circumvents mainline distribution channels that may incorporate secure policies
-of data integrity, authenticity, and privacy. Some piracy tools provide
-some ways to vet the authenticity of an uploader based on past uploads or based
-on the user's relationship with the piracy/torrenting
-site\footnote{%
-  \href{https://pirates-forum.org/Thread-ThePirateBay-Want-Trusted-VIP-Status-Pink-Green-Skull-See-Here}{ThePirateBay},
-  for example provides icons to display whether a user is \emph{Trusted},
-  \emph{VIP}, or \emph{Moderator}%
-}, but not all sites provide these metrics or warnings for the average end-user.
-Instead, the burden is often on the end-user, and it is not unfair to assume
-that the average person would not be able to analyze downloaded software for
-malware.
+of data integrity, authenticity, and privacy. There are piracy tools and sites
+that provide some ways to vet the authenticity of a user based on the user's
+previous history and relationship with the piracy community. For example,
+ThePirateBay, one of the most popular pirating sites, displays the status or
+class of their users and holds these users to higher standards in order to
+retain trust with the site's end users.~\cite{suprbay:status} Nevertheless, not
+all sites provide these metrics or warnings for the average end-user. Instead,
+the burden is often on the end-user, and it is not unfair to assume that the
+average person would not be able to analyze downloaded software for malware.
 
-% TODO: Add an example of malware distributed via piracy
+The relationship between piracy and malware distribution is not just a
+theoretical argument against the use of piracy. Although research has remained
+rather theoretical in this area, some researchers have performed empirical
+studies to find a stronger link between more prevalent use of piracy and
+malware presence. The study by Mezzour et al. reveals that piracy is not equal
+around the globe, and it poses serious problems especially for people in
+poor countries where piracy is widespread. Although they hypothesized, like
+other researchers, that \textcquote{mezzour:empirical-study}[,]{%
+  cyber criminals would target rich countries because the monetary benefit of
+  compromising rich people's computers is higher
+} but they found that, \textcquote{mezzour:empirical-study}[.]{%
+  surprisingly, poor countries in Sub-Saharan Africa because of the low cost of
+  attacking computers in this region%
+} Mezzour et al. conclude that, due to the high correlation between piracy
+and malware, one of the policy implications is that Sub-Saharan Africa requires
+reducing piracy to combat the spread of malware.
+
+% TODO: Add an example of malware distributed via piracy, see pcmag source
 
@@
 
-\textcquote{mrad:effects-of-ip}[.]{%
+\blockcquote{mrad:effects-of-ip}[.]{%
   According to Gould and Gruben (1996), IPRs protection stimulates economic
@@
 the debate. As shown above, piracy does have some merit to it. Christians have
-to admit that piracy-promoting tools do enable
+to admit that piracy-promoting tools do enable equal access for those in need,
+but, at the same time, we must recognize that the human nature is not perfect,
+but is full of evil thoughts (Matthew 15:19). Our human corruption manifests
+itself in many ways through our lives and actions, but, as explored in this
+paper, it also presents itself in the software we share. Our deceitful hearts
+(Jeremiah 17:9) make it hard to trust the software we download, but reckless
+use of piracy ignores advice and is therefore the way of the fool where people
+fall (Proverbs 12:15, 11:14).
 
-% 1 Peter 2:13-17
-% Romans 13:4-5
+On the one hand, a Christian use of piracy should be respectful of authority.
+We are called to do this regardless of punishment, but as a matter of
+conscience (Romans 13:4--5). We are to exercise our freedom respectfully and
+not as a cover-up for evil (1 Peter 2:13--7). In other words, we should consider
+both the consequences we may incur and the image we portray carefully before
+making a conclusion about the practice of piracy. A proper consideration of
+privacy should not ignore the serious ramifications and associations that come
+with it.
 
-\begin{enumerate}
-  \item Drawing distinctions: Paywall to information is different from
-    withholding information for sake of security
-
-  \item Paywalls provide the means of living for artists, writers, \&c. Providing
-    ways to circumvent these when unnecessary may hinder their livelihoods.
-
-  \item Betterment of mankind as a whole, should not be behind a paywall
-
-  \item Piracy provides an avenue for missionary outreach; piracy provides
-    means of secret dissemination within an oppressive regime.
-
-  \item Piracy provides the means for access to content/information without
-    much accountability
-\end{enumerate}
+On the other hand, piracy, when used properly, is able to be used for good
+causes that serve God rather than men (Acts 5:29). Academic articles,
+especially from medical journals, are crucial for the development of humanity
+and the saving of lives. In these cases, when capitalist motivators impede these
+greater goods, one should not turn a blind eye, but instead we should remember
+the poor and those in need (Proverbs 19:17, 22:9, 15:11; Galatians 2:10). Cases
+like these present valid concerns in which piracy appears to provide a viable
+solution to the problems brought about by human greed.
 

### 2024-04-10 paper.tex
finish conclusion

--- 2024-04-10 paper.tex
@@
 \section{Conclusion}
+In closing, piracy and its related tools are not as cut and dry of an issue as one might
+initially be led to believe. On one hand, piracy facilitates the spread of illegal goods
+and services. Pirated software from an untrusted source may also contain malware and the
+act of piracy itself, from one perspective, takes business and money away from companies.
+On the other hand, piracy can be a powerful force for good. It assists in the
+dissemination of information that is dangerous to oppressive regimes and it ensures that
+people of all financial backgrounds can access and contribute to valuable research for the
+benefit of everyone. More generally, the goal of piracy is to keep information free and
+open to everyone; in that regard, it promotes equality of access to human thoughts.
 
@@
 
-\begin{enumerate}
-  \item In response to economic argument: Would the pirate have bought the
-    product/service in the first place?
-\end{enumerate}
+By far, one of the most common arguments raised against piracy, as highlighted previously,
+is that piracy steals money from businesses because it demotivates people from paying for
+their products and services. While this may be true in many cases, the reality is a little
+more nuanced. Someone who chooses to pirate a product rather than purchase it may not have
+bought the product in the first place if piracy was not an option. On top of that, if
+someone really enjoys using the pirated version of a piece of software, they might choose
+to pay for the legitimate version of it to support the developers and get new updates
+faster. Even if the user refuses to pay for software, distributing the pirated version of
+software helps popularize it. When something gets really popular, people want to try it
+and that leads to more people who might decide to support the company who develops it by
+purchasing the real version.
 
-\begin{displayquote}
-  Of additional concern is the fact that by making \textins{Intellectual Property}
-  theft a national security threat without being clear about what actually
-  constitutes intellectual property, not only does the U.S. government create a
-  new reason for a militarized Internet, but it also sets the stage for companies
-  to assert that a range of other activities from file sharing to producing
-  counterfeit DVDs threaten national security and require further state
-  intervention.~\cite[p. 264]{halbert:agendas}
-\end{displayquote}
+Regardless of the potential benefit of piracy, businesses still see it as a clear threat
+to their bottom line and try their best to protect their intellectual property at the
+expense of the consumer's open access to information. \blockcquote[264]{halbert:agendas}
+{Of additional concern is the fact that by making \textins{Intellectual Property} theft a
+national security threat without being clear about what actually constitutes intellectual
+property, not only does the U.S. government create a new reason for a militarized
+Internet, but it also sets the stage for companies to assert that a range of other
+activities from file sharing to producing counterfeit DVDs threaten national security and
+require further state intervention.}
 
-\subsection{Applications}
-
-\subsection{Further Work}
+With how advanced piracy technology has become and how it will continue to improve, it
+will not be going away anytime soon. With that in mind, Christians should make an effort
+to seek the ways in which the tools of piracy enable the spread of the Gospel and how it
+can be used as a salt and a light to those in less fortunate places---fellow Christians
+and unbelievers alike.
 

### 2024-04-18 paper.tex
Start working on responses for Against sections

@@
 
+\subsubsection{Response}
+An argument that focuses solely on the criminality of the actions of piracy
+may present a strong deterrent for many that would consider piracy. It is true
+that the repercussions of piracy can be great, and it is true that in many
+cases both parties may be liable for damages or in danger of prison sentences.
+Christians should consider these things deeply. We are called to be meek,
+merciful, and peaceful (Matthew 5:5, 7, 9; Galatians 5:22-26). We are also
+called to harmony (Romans 12:16--18, Galatians 5:15-17). Christians have
+pondered on the questions of civil disobedience. While we are called to peace,
+love, and obedience, our obedience to man should not disrupt our obedience to
+God. Clearly, the Bible does not call for anarchy (see Romans 13), but it
+does present some cases in which civil disobedience for a greater good (that
+aligns with God's values) supersedes extreme patriotism.~\footnote{%
+  See the examples of the Hebrew midwives of Exodus 1; see also Rahab's defence
+  of the Israelite spies; see also Daniel's disobedience to King
+  Nebuchadnezzar; see also examples of civil disobedience in Revelation 13:15%
+} Christians have some precedent for civil disobedience, but, at the same time,
+Christians should also expect that there will be punishments for our actions.
+While the legislation against piracy is not always evil (and thus does not
+call for Christians to exert some civil disobedience), some of it can call our
+consciences into question, especially when we consider critical research such
+as the medical advances that have been discussed earlier in this paper.
+
 \subsubsection{Malware}
@@
 reducing piracy to combat the spread of malware.
+Antivirus provider Kaspersky recently uncovered an example of malware spreading
+through pirated software.~\cite{pcmag:malware}
 
-% TODO: Add an example of malware distributed via piracy, see pcmag source
+\subsubsection{Response}
+An argument based on malware distribution presents a strong case against
+piracy. Moreover, we have to acknowledge that the problem targets those with
+less resources (whether that be information, financial resources, or technical
+resources) to protect themselves from the threats that malware poses.
+
+% TODO
 
@@
 % TODO: Talk about ties to illegal organizations that are involved?
+
+\subsubsection{Response}
+
+% TODO
 
@@
 }
-Piracy, therefore, discourages innovation in the capitalist marketplace.
+Piracy, as proponents of this argument would contend, discourages innovation in
+the capitalist marketplace.
+
+\subsubsection{Response}
 
 % TODO
+
 \section{Christian Response}

### 2024-04-18 paper.tex
expand privacy section

--- 2024-04-10 paper.tex
@@
 \subsection{Privacy}
+It would be remiss to discuss the tools of piracy without mentioning virtual private
+networks (VPNs) and Tor, formerly known as the onion router. Commercially available VPNs
+allow internet users to route their traffic through another machine on the internet before
+connecting to the final destination host. What this means to the user is that their
+traffic appears to be coming from another IP address than the one for their home network.
+Since IP addresses are associated with geographical locations, a VPN can also mask a
+user's location from bad actors. However, VPNs are not a perfect solution to privacy. The
+company hosting the virtual private network is entrusted with all the user's internet
+traffic, so the company should not be left susceptible to cybersecurity attacks or
+coercion from government entities that might compromise someone's privacy online. In
+general, the user should have confidence in the VPN provider that their information will
+not be misused.
+
+In the context of piracy, a VPN enables users to bypass firewalls that block access to
+certain locations on the internet and they are employed in combination with the BitTorrent
+protocol to protect one's identity when they are connecting to seeders or leechers. When
+used responsibly, a privacy-conscious VPN provider is quite the boon to internet users in
+countries that oppress certain people groups. As long as the VPN itself is not
+compromised, a user's true identity can stay hidden. In the case of a missionary in a
+hostile environment, a VPN might enable them to bypass the restrictions that prevent them
+from accessing certain online editions of the Bible. Ultimately, VPNs have the ability to
+be a tool for good in situations where access to information is limited and where one
+might be persecuted if their identity were to be discovered on the internet.
+
+In the same vein as a VPN is Tor. Tor is a service that allows users to browse the
+internet without anyone but themselves knowing both the source and destination of a
+connection. Unlike a VPN, there is no central company that can see who is connecting to
+what destination; the only person who knows that is the one using the service. The Tor
+Project writes, \textcquote{tor:about}[.]{\textins{t}he goal of onion routing was to have
+a way to use the internet with as much privacy as possible, and the idea was to route
+traffic through multiple servers and encrypt it each step of the way} Tor not only
+provides privacy, it gives users total anonymity. Similar to using a VPN, connecting to
+the Tor network before browsing the internet makes it near impossible for observers to see
+what sites someone is accessing and for the site to see where the request is coming from.
+The potential applications for near-perfect anonymity on the internet are innumerable.
+Those living under oppressive governments can bypass censorship to reach the truth and
+Christians can anonymously access the Bible online in restrictive countries, just to name
+a few.
+
 As a side note, there are instances where pirated software can be more privacy-oriented

### 2024-04-18 paper.tex
Finish some responses

--- 2024-04-18 paper.tex
@@
 
-\title{Sailing the High Seas:\\ A Christian Exploration of Piracy}
+\title{Sailing the High Seas: A Christian Exploration of Piracy}
 \author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
@@
 \subsection{Background Information}
-
 \blockcquote{barlow:wine}{%
@@
 \subsection{Thesis/Argument}
-
 Piracy should not be hastily thrown aside.
@@
 
-\subsubsection{Illegal content}
+\subsubsection{Illegal and Immoral Content}
 Piracy, as discussed above, provides a way to distribute illegal content. This
@@
 the door for questionable and illegal content that may be harmful to society
-or national security.
+or national security. Since this paper aims to focus on a Christian exploration
+of piracy, then this section will focus mostly on immoral content to build a
+stronger argument against piracy that diversifies itself from an argument from
+criminality (as was explored earlier in this paper).
 
-% TODO: Talk about ties to illegal organizations that are involved?
+There are few (if any) types of illegal content that rival that of child
+pornography. It is a very sad reality, but this type of content is distributed
+widely through the use of the Internet. Why and how this happens is perhaps out
+of scope for this paper, but part of may be due to a sense of privacy and
+anonymity that is fostered to an extreme in certain subcultures of the
+Internet.~\cite[588]{prichard:subcultures} Research reveals that piracy and
+piracy-adjacent tools provide the means to distribute this disgusting content
+to unimaginable scales. Prichard et al explored this area of research, with
+the following conclusions:
+\blockcquote[593]{prichard:subcultures}{%
+  Based on longitudinal data collected between August and November 2010, this
+  study screened for the search terms that still appeared in the isoHunt Top
+  300 list in November after being observed during August. In summary, the
+  study found that 158/162 search terms related to infringing content, while
+  3/162 search terms related to child pornography and 1/162 to bestiality%
+}
+The consistently widespread nature of this immoral and illegal content presents
+a really compelling argument against piracy for Christians.
 
 \subsubsection{Response}
+As mentioned above, Christians should not quickly dismiss this argument against
+piracy. The reality of the matter is that we are sinful creatures full of
+evil thoughts (Matt. 15:9); we are in dire need of a savior to redeem us, not
+by any good works because even our best works are like filthy rags (Isa. 64:6).
+This argument is compelling because of its severity. A naive response would
+argue that harmful content is distributed via the surface web and without
+piracy anyways. The argument, however, holds in that piracy enables the
+widespread nature of this content, so what is a proper response to this
+argument? There may be ways to engage with piracy communities to self-regulate
+or to build better communities, as is suggested by Prichard et
+al.~\cite[595--596]{prichard:subcultures}
 
-% TODO
+This may be an unsatisfactory and complex answer to a large and complex issue.
+Piracy (namely in the form of torrenting) does, however, rely on peers to
+seed the content, and if changes can be made in these subcultures of the
+internet via evangelizing, perhaps there is hope for a better piracy community.
 
@@
 \subsubsection{Response}
+While this argument does hold in many cases, the economic reality is more
+complex. The argument assumes that pirates would pay for the content if piracy
+was not an option, but that assumption should be brought into question. As
+argued earlier in this paper, piracy can be a means to control the market for
+real change, and pirates maybe do not consider purchasing the content in the
+first place.
 
-% TODO
+On the other hand, it is not completely out of the question to imagine a
+pirate coming back to some creator (whether that is an individual or a
+business) to pay for the content when they are finally able to. As this paper
+showed earlier, many pirates are those who do not have the means to pay for
+some of the content that they pirate, but perhaps they would if/when they
+could.
 
@@
 \subsection{Final Thoughts / Responses}
-
 By far, one of the most common arguments raised against piracy, as highlighted previously,

### 2024-04-18 paper.tex
add response to malware

--- 2024-04-18 paper.tex
@@
 Piracy may openly distribute dangerous illegal content, but arguably the
-larger threat is the dangerous content that remains undisclosed. Piracy
-circumvents mainline distribution channels that may incorporate secure policies
-of data integrity, authenticity, and privacy. There are piracy tools and sites
-that provide some ways to vet the authenticity of a user based on the user's
-previous history and relationship with the piracy community. For example,
-ThePirateBay, one of the most popular pirating sites, displays the status or
-class of their users and holds these users to higher standards in order to
-retain trust with the site's end users.~\cite{suprbay:status} Nevertheless, not
-all sites provide these metrics or warnings for the average end-user. Instead,
-the burden is often on the end-user, and it is not unfair to assume that the
-average person would not be able to analyze downloaded software for malware.
-
+larger threat is the dangerous content that remains undisclosed. 
 The relationship between piracy and malware distribution is not just a
@@
 \subsubsection{Response}
-An argument based on malware distribution presents a strong case against
-piracy. Moreover, we have to acknowledge that the problem targets those with
-less resources (whether that be information, financial resources, or technical
-resources) to protect themselves from the threats that malware poses.
-
-% TODO
+An argument based on malware distribution presents a strong case against piracy. Moreover,
+we have to acknowledge that the problem targets those with less resources (whether that be
+information, financial resources, or technical resources) to protect themselves from the
+threats that malware poses. Piracy circumvents mainline distribution channels that may
+incorporate secure policies of data integrity, authenticity, and privacy. However, there
+are piracy tools and sites that provide some ways to vet the authenticity of a user based
+on the user's previous history and relationship with the piracy community. For example,
+ThePirateBay, one of the most popular pirating sites, displays the status or class of
+their users and holds these users to higher standards in order to retain trust with the
+site's end users.~\cite{suprbay:status} Nevertheless, not all sites provide these metrics
+or warnings for the average end-user. Instead, the burden is often on the end-user, and
+it is not unfair to assume that the average person would not be able to analyze downloaded
+software for malware. Thus, to safely make use of piracy tools, one must be aware of the
+risks and the ways that they can mitigate those. This is not to say that users need to
+reverse engineer all the software they download to look for malicious behavior, but they
+do need to be aware of software sources that give them a much better shot at getting an
+untainted copy. As mentioned previously, user trust is a crucial component of piracy;
+those who choose to enter this domain should be careful to download only from those with a
+high reputation in the community---backed by a long time of faithful service.
 

### 2024-04-24 paper.tex
Fix economic section

@@
 \subsection{Economic}
-Piracy, as a theft, also causes real harm to businesses. The act of piracy
-does involve stealing and distributing content from creators and organizations
-that invested time, money, and effort into the content. In this sense, piracy
-is stealing from the individuals' revenue and the government's tax
-revenue.~\cite{congress:pirating-the-american-dream}
+Piracy, as a form of theft, is also able to cause real harm to businesses. The
+act of piracy can and does involve stealing and distributing content from
+creators and organizations that invested time, money, and effort into the
+content. In this sense, piracy is stealing from the individuals' revenue and
+the government's tax revenue.~\cite{congress:pirating-the-american-dream}
 
 \blockcquote{mrad:effects-of-ip}[.]{%
-  According to Gould and Gruben (1996), IPRs protection stimulates economic
+  According to Gould and Gruben (1996),
+  \textins{Intellectual Property Rights (IPRs)} protection stimulates economic
   growth if it is accompanied by a policy of trade liberalization. By
@@
 Piracy, as proponents of this argument would contend, discourages innovation in
-the capitalist marketplace.
+the capitalist marketplace. Conversely, they argue that protecting intellectual
+property leads to growth and innovation. Piracy, then, undermines the means for
+economic growth that would benefit the general population.
 

### 2024-04-24 paper.tex
Add Christian perspective to Response sections

@@
 \subsubsection{Response}
-An argument based on malware distribution presents a strong case against piracy. Moreover,
-we have to acknowledge that the problem targets those with less resources (whether that be
-information, financial resources, or technical resources) to protect themselves from the
-threats that malware poses. Piracy circumvents mainline distribution channels that may
-incorporate secure policies of data integrity, authenticity, and privacy. However, there
-are piracy tools and sites that provide some ways to vet the authenticity of a user based
-on the user's previous history and relationship with the piracy community. For example,
-ThePirateBay, one of the most popular pirating sites, displays the status or class of
-their users and holds these users to higher standards in order to retain trust with the
-site's end users.~\cite{suprbay:status} Nevertheless, not all sites provide these metrics
-or warnings for the average end-user. Instead, the burden is often on the end-user, and
-it is not unfair to assume that the average person would not be able to analyze downloaded
-software for malware. Thus, to safely make use of piracy tools, one must be aware of the
-risks and the ways that they can mitigate those. This is not to say that users need to
-reverse engineer all the software they download to look for malicious behavior, but they
-do need to be aware of software sources that give them a much better shot at getting an
-untainted copy. As mentioned previously, user trust is a crucial component of piracy;
-those who choose to enter this domain should be careful to download only from those with a
-high reputation in the community---backed by a long time of faithful service.
+An argument based on malware distribution presents a strong case against piracy.
+Moreover, we have to acknowledge that the problem targets those with less
+resources (whether that be information, financial resources, or technical
+resources) to protect themselves from the threats that malware poses. Piracy
+circumvents mainline distribution channels that may incorporate secure policies
+of data integrity, authenticity, and privacy. However, there are piracy tools
+and sites that provide some ways to vet the authenticity of a user based on the
+user's previous history and relationship with the piracy community. For example,
+ThePirateBay, one of the most popular pirating sites, displays the status or
+class of their users and holds these users to higher standards in order to
+retain trust with the site's end users.~\cite{suprbay:status}
+
+Nevertheless, not all sites provide these metrics or warnings for the average
+end-user. Instead, the burden is often on the end-user, and it is not unfair to
+assume that the average person would not be able to analyze downloaded software
+for malware. Thus, to safely make use of piracy tools, one must be aware of the
+risks and the ways that they can mitigate those. This is not to say that users
+need to reverse engineer all the software they download to look for malicious
+behavior, but they do need to be aware of software sources that give them a
+much better shot at getting an untainted copy. As mentioned previously, user
+trust is a crucial component of piracy; those who choose to enter this domain
+should be careful to download only from those with a high reputation in the
+community---backed by a long time of faithful service.
+
+The issue of trust is one that should not be unfamiliar to Christians. As
+Christians, we have a Lord that we can trust in with all our hearts (Prov. 3:5).
+As the Psalmist writes, \enquote{It is better to take refuge in the Lord than
+to trust in man} (Ps. 118:8). Christians should be slow to put their entire
+trust in metrics like those provided by ThePirateBay; while they provide some
+founts of trust, they should not immediately encourage Christians to trust
+piracy.
 
@@
 some of the content that they pirate, but perhaps they would if/when they
-could.
+could. This possibility of turnaround should be the focus of further research
+in the culture of piracy, as the economic argument (for either side) largely
+depends on whether pirates would pay for content if they could.
+
+While economic growth can and does benefit the general population, a Christian
+response to piracy should take into consideration some of the more complex
+aspects of the global marketplace. It is true that the growth of a business can
+benefit the employees generally, but the well-being of the consumer should also
+be factored into the equation. In the case of commodity software such as
+movies, music, or games, the problem that leads to widespread piracy in some
+regions is an economic problem; the prices for software or media is set for and
+targeted towards an affluent audience. If the companies do not set regional
+prices that make their offerings reasonably available to a population, then it
+should not come as a surprise that piracy will be widespread. As Christians and
+computer scientists, we should be generous to the poor and set regional prices
+that make sense for consumers if piracy is negatively affecting the people
+(e.g., employees) that we are responsible for (Prov. 14:31; 19:17; Lk. 12:33-34).
 

### 2024-04-24 paper.tex
phrasing
Distinguish bibliography from references
Implement some recommended edits

@@
 \maketitle
+
+\begin{refsection}
 
@@
 \subsection{Thesis/Argument}
-Piracy should not be hastily thrown aside.
-Piracy is an excellent tool for disseminating information, services, content,
-\&c. that may not have been available any other way.
+Piracy should not be hastily thrown aside nor should it be accepted
+wholeheartedly in every situation. Piracy is an excellent tool for
+disseminating information, services, content, \&c. that may not have been
+available any other way. As a Christian, a proper exploration of piracy should
+consider the potential benefits (or circumstances that warrant piracy) and the
+potential downsides (or circumstances that need special attention from
+Christian developers, writers, and netizens).
 
@@
 equal access to information, governance, and privacy.
+
 \subsection{Equal Access}
@@
 Christians should also expect that there will be punishments for our actions.
-While the legislation against piracy is not always evil (and thus does not
-call for Christians to exert some civil disobedience), some of it can call our
-consciences into question, especially when we consider critical research such
-as the medical advances that have been discussed earlier in this paper.
+
+As the arguments against piracy that follow in this paper, one should consider
+that legislation against piracy is not always evil; there are valid concerns
+with piracy's immoral/illegal content and economic damages. Christians should
+be careful to call upon civil disobedience as a defense for their piracy habits
+online. On the other hand, we also need to consider some cases in which being
+a seeder could benefit our neighbors with less resources to legally obtain
+research, software, or files that could benefit the general population.
 
@@
 poor countries where piracy is widespread. Although they hypothesized, like
-other researchers, that \textcquote{mezzour:empirical-study}[,]{%
-  cyber criminals would target rich countries because the monetary benefit of
-  compromising rich people's computers is higher
-} but they found that, \textcquote{mezzour:empirical-study}[.]{%
-  surprisingly, poor countries in Sub-Saharan Africa because of the low cost of
-  attacking computers in this region%
+other researchers, that \enquote{cyber criminals would target rich countries}
+because of monetary incentives, they found that, cyber criminals targeted
+\textcquote{mezzour:empirical-study}[.]{%
+  poor countries in Sub-Saharan Africa because of the low cost of attacking
+  computers in this region%
 } Mezzour et al. conclude that, due to the high correlation between piracy
@@
 reducing piracy to combat the spread of malware.
-Antivirus provider Kaspersky recently uncovered an example of malware spreading
-through pirated software.~\cite{pcmag:malware}
+In more recent news, Antivirus provider Kaspersky recently uncovered an example
+of malware spreading through pirated software.~\cite{pcmag:malware}
 
@@
 
-There are few (if any) types of illegal content that rival that of child
-pornography. It is a very sad reality, but this type of content is distributed
-widely through the use of the Internet. Why and how this happens is perhaps out
-of scope for this paper, but part of may be due to a sense of privacy and
-anonymity that is fostered to an extreme in certain subcultures of the
-Internet.~\cite[588]{prichard:subcultures} Research reveals that piracy and
+There are few (if any) types of illegal content that are as immoral as that of
+child pornography. It is a very sad reality, but this type of content is
+distributed widely through the use of the Internet. Why and how this happens is
+perhaps out of scope for this paper, but part of may be due to a sense of
+privacy and anonymity that is fostered to an extreme in certain subcultures of
+the Internet.~\cite[588]{prichard:subcultures} Research reveals that piracy and
 piracy-adjacent tools provide the means to distribute this disgusting content
@@
 }
-The consistently widespread nature of this immoral and illegal content presents
-a really compelling argument against piracy for Christians.
+These numbers may seem small to many. To properly interpret the above numbers,
+consider that: 1) they do not represent the quantity of files that are shared
+and 2) the researchers focused on isoHunt (which operated on the surface web)
+and did not explore other means to find torrent files (i.e., means that utilize
+the anonymity of the dark web). The numbers, rather, represent the consistent
+interest within certain online subcultures in the immoral content. This
+consistency presents a really compelling argument for Christians to question
+piracy and the online culture that promotes and uses these methods.
 
@@
 seed the content, and if changes can be made in these subcultures of the
-internet via evangelizing, perhaps there is hope for a better piracy community.
+internet. Because of the (mostly) decentralized nature of piracy, the best hope
+is for the popular sites and individuals to self-regulate the content that is
+promoted or accepted. For Christians, one option to keep in mind or consider
+is to evangelize or set an example of what good online behavior looks like.
+These questions of regulation and changes in the online culture, however,
+reside outside the scope of this paper, but they, nevertheless, remain (and
+will remain) increasingly relevant.
 
@@
 making a conclusion about the practice of piracy. A proper consideration of
-privacy should not ignore the serious ramifications and associations that come
+piracy should not ignore the serious ramifications and associations that come
 with it.
@@
 \clearpage
+\printbibliography[title=References]
+\end{refsection}
+
+\clearpage
 \nocite{*}
-\printbibliography
+\printbibliography[title=Bibliography]
 \end{document}

### 2024-04-24 paper.tex
Edit packages used.
Also add epigraph package for formatting

@@
 \usepackage{csquotes}
+\usepackage{epigraph}
 \usepackage{footnote}
-\usepackage{endnotes}
 \usepackage[bottom, multiple]{footmisc}
-\usepackage[level]{fmtcount}
-
-\usepackage{float}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\graphicspath{ {./graphics} }
 
@@
 
-\makesavenoteenv{figure}
 \SetCiteCommand{\cite}
@@
 \begin{document}
+\begin{refsection}
 \maketitle
-
-\begin{refsection}
-
-\section{Introduction}
-
-\subsection{Background Information}
-\blockcquote{barlow:wine}{%
+\epigraph{%
   \textins{T}hat ideas should freely spread from one to another over the globe,
@@
   have our physical being, incapable of confinement or exclusive appropriation.
-  Inventions then cannot, in nature, be a subject of property.
-  \emph{Thomas Jefferson}%
-}
+  Inventions then cannot, in nature, be a subject of property.%
+}{Thomas Jefferson~\cite{barlow:wine}}
 
+\section{Introduction}
+\subsection{Background Information}
 It is generally accepted that piracy is an illegal activity, but what is often overlooked

### 2024-04-28 paper.tex
Edit titling
Trailing whitespaces
formatting of .tex file

@@
 \usepackage{setspace}
+\usepackage{titling}
+\usepackage{authblk}
 
@@
 \addbibresource{sources.bib}
-
-\usepackage[pdfusetitle, hidelinks]{hyperref}
-
 \SetCiteCommand{\cite}
 
+\usepackage[hidelinks]{hyperref}
+
+\makeatletter
 \title{Sailing the High Seas: A Christian Exploration of Piracy}
-\author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
-\date{Wheaton College\\\today}
+\author{Pedro Avalos Jim\'enez}
+\author{Niklas Anderson}
+\affil{Wheaton College}
+\date{\today}
+\makeatother
 
@@
 with which ideas propagate lead people like Thomas Jefferson to claim that they are
-``incapable of confinement or exclusive appropriation.'' Consequently, two parties form:
+\enquote{incapable of confinement or exclusive appropriation.} Consequently, two parties form:
 those that support piracy for its free and open spread of all information and those that
@@
 what piracy means insofar as how this paper interprets it. The dictionary defines piracy
-as ``the unauthorized use of another's production, invention, or conception especially in
-infringement of a copyright.'' While this covers the technical definition of piracy, this
+as \enquote{the unauthorized use of another's production, invention, or conception especially in
+infringement of a copyright.} While this covers the technical definition of piracy, this
 paper elects to utilize a more wholistic definition. Piracy not only includes the
@@
 upon the research that has already been done. John Barlow puts the situation eloquently in
-\textit{Selling Wine Without Bottles: The Economy of the Mind on the Global Net} when he
-says, \textcquote{barlow:wine}{I am not comfortable with a model which will restrict
-inquiry to the wealthy.} In the case of academic journal articles, why shouldn't anyone be
-able to better themselves by consuming quality research articles? All humans should be
-allowed to continue to learn more about God's creation without being bound by their
-financial status. Just as how someone can go to the library and checkout any book that
-tickles their fancy, they should be freely able to learn from the ongoing research of our
-world. Of course one should recognize that not all information is safe or helpful to be
-freely available on the internet, but that is a topic for a later section.
+\citetitle{barlow:wine} when he says, \textcquote{barlow:wine}[.]{%
+  I am not comfortable with a model which will restrict inquiry to the wealthy%
+}
+In the case of academic journal articles, why shouldn't anyone be able to better
+themselves by consuming quality research articles? All humans should be allowed
+to continue to learn more about God's creation without being bound by their
+financial status. Just as how someone can go to the library and checkout any
+book that tickles their fancy, they should be freely able to learn from the
+ongoing research of our world. Of course one should recognize that not all
+information is safe or helpful to be freely available on the internet, but that
+is a topic for a later section.
 
@@
 with regards to medical research. Till et al. stresses how important access to medical
-research is in their article \textit{Who is pirating medical literature? A bibliometric
-review of 28 million Sci-Hub downloads}. \blockcquote{till:medical-literature} {Access to
-the medical literature is essential for both the practice of evidence-based medicine and
-meaningful contribution to medical sciences. Nonetheless, only 12\% of newly published
-papers are freely accessible online, and, as of 2014, only 3 million of the 26.3 million
-articles indexed on PubMed were available on the site's repository of free materials,
-PubMed Central. Access to paywall-protected literature remains primarily through
-institutional subscriptions. Such subscriptions are costly and many struggle to afford
-access. The result is a disparity in access to the medical literature, particularly for
-those in low-income and middle-income countries (LMICs).}
+research is in their article \citetitle{till:medical-literature}.
+\blockcquote{till:medical-literature}[.]{%
+  Access to the medical literature is essential for both the practice of
+  evidence-based medicine and meaningful contribution to medical sciences.
+  Nonetheless, only 12\% of newly published papers are freely accessible online,
+  and, as of 2014, only 3 million of the 26.3 million articles indexed on PubMed
+  were available on the site's repository of free materials, PubMed Central.
+  Access to paywall-protected literature remains primarily through institutional
+  subscriptions. Such subscriptions are costly and many struggle to afford
+  access. The result is a disparity in access to the medical literature,
+  particularly for those in low-income and middle-income countries (LMICs)%
+}
 According to Till et al. the high subscription cost to access medical journal articles
@@
 is detrimental to the furthering of medical sciences. In light of this, it should come as
-no surprise that \textcquote{till:medical-literature} {Nearly 1 million articles published
-by medical journals are downloaded on Sci-Hub each month.} Piracy enables the less
-fortunate countries to access valuable medical data that is essential to furthering
-medical research for the human race. In such a specific case as this, it should not be
-controversial to conclude that piracy is a good thing.
+no surprise that \textcquote{till:medical-literature}[.]{
+  Nearly 1 million articles published by medical journals are downloaded on
+  Sci-Hub each month%
+} Piracy enables the less fortunate countries to access valuable medical data that is
+essential to furthering medical research for the human race. In such a specific case as
+this, it should not be controversial to conclude that piracy is a good thing.
 
@@
 control. The government's imposing of restrictions on the Internet is clearly seen when
-\textcquote{current:jigsaw} {Tunisia significantly ramped up its already aggressive
-blocking of specific websites in response to unrest that would ultimately unseat its
-government.} Even more extreme is when governments shutdown the internet entirely to
-quickly cease the spread of information. This is exactly what happened on January 28, 2011
-in Egypt: \textcquote{current:jigsaw} {Almost simultaneously, about 3,500 individual
-Border Gateway Protocol routes \textelp{} were withdrawn on orders from the Egyptian
-government, cutting the country off from the rest of the world and bringing internal
-communication to a halt.} Fortunately, as already mentioned earlier, piracy excels at
-spreading information even when there are those that do not want it spread. BitTorrent is
+\textcquote{current:jigsaw}[.]{%
+  Tunisia significantly ramped up its already aggressive blocking of specific
+  websites in response to unrest that would ultimately unseat its government%
+} Even more extreme is when governments shutdown the internet entirely to
+quickly cease the spread of information. This is exactly what happened on
+January 28, 2011 in Egypt: \textcquote{current:jigsaw}[.]{%
+  Almost simultaneously, about 3,500 individual Border Gateway Protocol routes
+  \textelp{} were withdrawn on orders from the Egyptian government, cutting the
+  country off from the rest of the world and bringing internal communication to
+  a halt%
+} Fortunately, as already mentioned earlier, piracy excels at spreading
+information even when there are those that do not want it spread. BitTorrent is
 a peer-to-peer file sharing protocol that is often used when pirating content on the
@@
 seek to do: encourage innovation. Mark Lemley presents an argument against IP laws when he
-writes, \blockcquote[1339]{lemley:faith-based}{[IP] intervenes in the market to interfere
-with the freedom of others to do what they want in hopes of achieving the end of
-encouraging creativity. If we take that purpose out of the equation, we are left with a
-belief system that says the government should restrict your speech and freedom of action
-in favor of mine, not because doing so will improve the world, but simply because I spoke
-first.} 
+writes, \blockcquote[1339]{lemley:faith-based}[.]{%
+  \textins{IP} intervenes in the market to interfere with the freedom of others
+  to do what they want in hopes of achieving the end of encouraging creativity.
+  If we take that purpose out of the equation, we are left with a belief system
+  that says the government should restrict your speech and freedom of action in
+  favor of mine, not because doing so will improve the world, but simply because
+  I spoke first%
+}
 To put it differently, intellectual property laws are based on the assumption that in
@@
 what destination; the only person who knows that is the one using the service. The Tor
-Project writes, \textcquote{tor:about}[.]{\textins{t}he goal of onion routing was to have
-a way to use the internet with as much privacy as possible, and the idea was to route
-traffic through multiple servers and encrypt it each step of the way} Tor not only
-provides privacy, it gives users total anonymity. Similar to using a VPN, connecting to
-the Tor network before browsing the internet makes it near impossible for observers to see
-what sites someone is accessing and for the site to see where the request is coming from.
-The potential applications for near-perfect anonymity on the internet are innumerable.
-Those living under oppressive governments can bypass censorship to reach the truth and
-Christians can anonymously access the Bible online in restrictive countries, just to name
-a few.
+Project writes, \textcquote{tor:about}[.]{%
+  \textins{t}he goal of onion routing was to have a way to use the internet with
+  as much privacy as possible, and the idea was to route traffic through
+  multiple servers and encrypt it each step of the way%
+} Tor not only provides privacy, it gives users total anonymity. Similar to
+using a VPN, connecting to the Tor network before browsing the internet makes it
+near impossible for observers to see what sites someone is accessing and for the
+site to see where the request is coming from.  The potential applications for
+near-perfect anonymity on the internet are innumerable. Those living under
+oppressive governments can bypass censorship to reach the truth and Christians
+can anonymously access the Bible online in restrictive countries, just to name a
+few.
 
 As a side note, there are instances where pirated software can be more privacy-oriented
-than the paid version. Sometimes the pirated version is ``cracked'' to remove certain
+than the paid version. Sometimes the pirated version is \emph{cracked} to remove certain
 digital rights management (DRM) tools that often run background tasks unbeknownst to the
@@
 of piracy, at least in the United States or nations where the individual
-downloading can be found guilty of a crime. ``Fair use'' and statutes of
+downloading can be found guilty of a crime. \emph{Fair use} and statutes of
 limitations may alleviate penalties, but repeat offences may also lose one's
@@
 to their bottom line and try their best to protect their intellectual property at the
-expense of the consumer's open access to information. \blockcquote[264]{halbert:agendas}
-{Of additional concern is the fact that by making \textins{Intellectual Property} theft a
-national security threat without being clear about what actually constitutes intellectual
-property, not only does the U.S. government create a new reason for a militarized
-Internet, but it also sets the stage for companies to assert that a range of other
-activities from file sharing to producing counterfeit DVDs threaten national security and
-require further state intervention.}
+expense of the consumer's open access to information.
+\blockcquote[264]{halbert:agendas}[.]{%
+  Of additional concern is the fact that by making
+  \textins{Intellectual Property} theft a national security threat without being
+  clear about what actually constitutes intellectual property, not only does the
+  U.S. government create a new reason for a militarized Internet, but it also
+  sets the stage for companies to assert that a range of other activities from
+  file sharing to producing counterfeit DVDs threaten national security and
+  require further state intervention%
+}
 

### 2024-04-28 paper.tex
Add abstract

@@
 \maketitle
+
 \epigraph{%
@@
 }{Thomas Jefferson~\cite{barlow:wine}}
+
+\begin{abstract}
+  This study is a comprehensive study of piracy, particularly in the digital
+  realm, from a Christian perspective. The paper delves into the ethical, legal,
+  and moral implications of piracy, acknowledging both its potential benefits
+  and drawbacks. The authors argue that piracy can promote equal access to
+  information, challenge oppressive regimes, and protect privacy. However, they
+  also highlight the risks associated with piracy, such as the distribution of
+  dangerous or illegal content and its negative economic impact. The paper
+  concludes with a call for a nuanced Christian response to piracy, emphasizing
+  the importance of respecting authority and considering the potential
+  consequences of piracy while considering the staying power and benefits of
+  piracy-promoting tools and practices. This work encourages a thoughtful
+  exploration of piracy, rather than a hasty dismissal or acceptance.
+\end{abstract}
 

### 2024-04-28 paper.tex
Add metadata for generated pdf.
Use citeauthor

@@
 \date{\today}
+
+\hypersetup{
+  pdftitle={\@title},
+  pdfauthor={Pedro Avalos Jim\'enez and Niklas Anderson},
+  pdfsubject={piracy},
+  pdfkeywords={%
+    Christian, piracy, benefits, drawbacks, exploration, nuance, governance,
+    privacy, malware, equal access, criminality, illegal content, economy%
+  },
+}
 \makeatother
@@
 high cost of entry, certain groups are automatically excluded from reading and building
-upon the research that has already been done. John Barlow puts the situation eloquently in
-\citetitle{barlow:wine} when he says, \textcquote{barlow:wine}[.]{%
+upon the research that has already been done. \citeauthor{barlow:wine} puts the situation
+eloquently in \citetitle{barlow:wine} when he says, \textcquote{barlow:wine}[.]{%
   I am not comfortable with a model which will restrict inquiry to the wealthy%
@@
 A more concrete instance of where paywalls can be directly harmful to the human race is
-with regards to medical research. Till et al. stresses how important access to medical
-research is in their article \citetitle{till:medical-literature}.
+with regards to medical research. \citeauthor{till:medical-literature} stresses how important
+access to medical research is in their article \citetitle{till:medical-literature}.
 \blockcquote{till:medical-literature}[.]{%
@@
 }
-According to Till et al. the high subscription cost to access medical journal articles
-means that those in less advantaged countries cannot afford access to the material, which
-is detrimental to the furthering of medical sciences. In light of this, it should come as
-no surprise that \textcquote{till:medical-literature}[.]{
+According to \citeauthor{till:medical-literature} the high subscription cost to access medical
+journal articles means that those in less advantaged countries cannot afford access to the
+material, which is detrimental to the furthering of medical sciences. In light of this, it
+should come as no surprise that \textcquote{till:medical-literature}[.]{
   Nearly 1 million articles published by medical journals are downloaded on
@@
 one another. That being said, it is worth examining whether IP laws accomplish what they
-seek to do: encourage innovation. Mark Lemley presents an argument against IP laws when he
-writes, \blockcquote[1339]{lemley:faith-based}[.]{%
+seek to do: encourage innovation. \citeauthor{lemley:faith-based} presents an argument
+against IP laws when he writes, \blockcquote[1339]{lemley:faith-based}[.]{%
   \textins{IP} intervenes in the market to interfere with the freedom of others
@@
 \subsubsection{Malware}
-The intersection of trust and computing is not a new topic. Ken Thompson, the
-creator of the Unix operating system, made this the topic of his Turing Award
-lecture. In his lecture, he poses the question,
+The intersection of trust and computing is not a new topic.
+\citeauthor{thompson:trust}, the creator of the Unix operating system, made this
+the topic of his Turing Award lecture. In his lecture, he poses the question,
 \textcquote[761]{thompson:trust}[.]{%
@@
 } With piracy, however, we add an additional actor in this chain of trust.
-Thompson's lecture focuses solely on the developer-to-user relationship of
-trust, with the assumption of an open-source project. Thompson, thus, comments
-on how \textcquote[763]{thompson:trust}{%
+\citeauthor{thompson:trust} focuses solely on the developer-to-user relationship
+of trust, with the assumption of an open-source project.
+\citeauthor{thompson:trust}, thus, comments on how
+\textcquote[763]{thompson:trust}{%
   \textins{s}uch blatant code would not go undetected for long. Even the most
@@
 studies to find a stronger link between more prevalent use of piracy and
-malware presence. The study by Mezzour et al. reveals that piracy is not equal
-around the globe, and it poses serious problems especially for people in
-poor countries where piracy is widespread. Although they hypothesized, like
-other researchers, that \enquote{cyber criminals would target rich countries}
-because of monetary incentives, they found that, cyber criminals targeted
-\textcquote{mezzour:empirical-study}[.]{%
+malware presence. The study by \citeauthor{mezzour:empirical-study} reveals that
+piracy is not equal around the globe, and it poses serious problems especially
+for people in poor countries where piracy is widespread. Although they
+hypothesized, like other researchers, that \enquote{cyber criminals would target
+rich countries} because of monetary incentives, they found that, cyber criminals
+targeted \textcquote{mezzour:empirical-study}[.]{%
   poor countries in Sub-Saharan Africa because of the low cost of attacking
   computers in this region%
-} Mezzour et al. conclude that, due to the high correlation between piracy
-and malware, one of the policy implications is that Sub-Saharan Africa requires
-reducing piracy to combat the spread of malware.
+} \citeauthor{mezzour:empirical-study} concludes that, due to the high
+correlation between piracy and malware, one of the policy implications is that
+Sub-Saharan Africa requires reducing piracy to combat the spread of malware.
 In more recent news, Antivirus provider Kaspersky recently uncovered an example
@@
 user's previous history and relationship with the piracy community. For example,
-ThePirateBay, one of the most popular pirating sites, displays the status or
-class of their users and holds these users to higher standards in order to
+\emph{ThePirateBay}, one of the most popular pirating sites, displays the status
+or class of their users and holds these users to higher standards in order to
 retain trust with the site's end users.~\cite{suprbay:status}
@@
 to trust in man} (Ps. 118:8). Christians should be slow to put their entire
-trust in metrics like those provided by ThePirateBay; while they provide some
-founts of trust, they should not immediately encourage Christians to trust
+trust in metrics like those provided by \emph{ThePirateBay}; while they provide
+some founts of trust, they should not immediately encourage Christians to trust
 piracy.
@@
 piracy-adjacent tools provide the means to distribute this disgusting content
-to unimaginable scales. Prichard et al explored this area of research, with
-the following conclusions:
+to unimaginable scales. \citeauthor{prichard:subcultures} explored this area of
+research, with the following conclusions:
 \blockcquote[593]{prichard:subcultures}{%
@@
 argument? There may be ways to engage with piracy communities to self-regulate
-or to build better communities, as is suggested by Prichard et
-al.~\cite[595--596]{prichard:subcultures}
+or to build better communities, as is suggested by
+\citeauthor{prichard:subcultures}.~\cite[595--596]{prichard:subcultures}
 

### 2024-04-29 paper.tex
add Christian connection to free access
add sources to unsubstantiated section

@@
 In the case of academic journal articles, why shouldn't anyone be able to better
-themselves by consuming quality research articles? All humans should be allowed
-to continue to learn more about God's creation without being bound by their
-financial status. Just as how someone can go to the library and checkout any
-book that tickles their fancy, they should be freely able to learn from the
-ongoing research of our world. Of course one should recognize that not all
-information is safe or helpful to be freely available on the internet, but that
-is a topic for a later section.
+themselves by consuming quality research articles? All humans should be allowed to
+continue to learn more about God's creation without being bound by their financial status.
+The Bible encourages one to pursue knowledge. Proverbs 1:5 reads, \textquote{\textins{l}et
+the wise hear and increase in learning, and the one who understands obtain guidance.}
+One's ability to live out this verse should not be artificially hindered by financial
+station. Just as how someone can go to the library and checkout any book that tickles
+their fancy, they should be freely able to learn from the ongoing research of our world.
+Of course one should recognize that not all information is safe or helpful to be freely
+available on the internet, but that is a topic for a later section.
 
@@
 cannot afford it, they might start to look for alternative way of obtaining it. But, when
-they can afford it and they think it is fairly priced, they will purchase it. Thus, piracy
-is often where tech-savvy persons turn to when they need something that they feel is
-exorbitantly priced, but is not often utilized when they feel the purchase is worthwhile.
-This way, piracy acts as a safeguard for consumers when they feel they are being taken
-advantage of by predatory business practices.
-% XXX the statements in the paragraph above are largely unsubstantiated >:D
-% XXX add darnton:pirating-and-publishing and bohannon:everyone sources if needed
+they can afford it and they think it is fairly priced, they will purchase it.
+
+According to a survey conducted by MUSO, \textcquote{muso-about}{a data company that
+provides the most complete and trusted view of global piracy and unlicensed media
+consumption,} the most common reasons that users pirate is \textcquote{vice-pirates}{the
+content they were looking for wasn’t legally available (34 percent) was too cumbersome or
+difficult to access (34 percent), or wasn’t affordable (35 percent).} The reason most
+people pirate is not that they are not willing to pay for the content, but that they
+cannot legally obtain it any other way, that it was too hard to access, or that it was not
+affordable. In the latter two cases, one can see the consumer's power to protest poorly
+designed systems or unaffordable prices. Thus, piracy is often where tech-savvy persons
+turn to when they need something that they feel is exorbitantly priced, but is not often
+utilized when they feel the purchase is worthwhile. This way, piracy acts as a safeguard
+for consumers when they feel they are being taken advantage of by predatory business
+practices.
 

### 2024-04-29 paper.tex
Formatting and add dictionary source

--- 2024-04-28 paper.tex
@@
 with which ideas propagate lead people like Thomas Jefferson to claim that they are
-\enquote{incapable of confinement or exclusive appropriation.} Consequently, two parties form:
-those that support piracy for its free and open spread of all information and those that
-see piracy as a threat to the further advancement of technology and ideas.
+\textcquote{barlow:wine}[.]{incapable of confinement or exclusive appropriation}
+Consequently, two parties form: those that support piracy for its free and open spread
+of all information and those that see piracy as a threat to the further advancement of
+technology and ideas.
 
@@
 what piracy means insofar as how this paper interprets it. The dictionary defines piracy
-as \enquote{the unauthorized use of another's production, invention, or conception especially in
-infringement of a copyright.} While this covers the technical definition of piracy, this
-paper elects to utilize a more wholistic definition. Piracy not only includes the
-unauthorized use of someone's work, but it includes the tools and techniques that are
-commonly used to accomplish this task such as torrent trackers, VPNs, and torrent clients.
+as \textcquote{merriam-webster:piracy}[.]{%
+  the unauthorized use of another's production, invention, or conception
+  especially in infringement of a copyright%
+} While this covers the technical definition of piracy, this paper elects to utilize a
+more wholistic definition. Piracy not only includes the unauthorized use of someone's
+work, but it includes the tools and techniques that are commonly used to accomplish
+this task such as torrent trackers, VPNs, and torrent clients.
 
@@
 In the case of academic journal articles, why shouldn't anyone be able to better
-themselves by consuming quality research articles? All humans should be allowed
-to continue to learn more about God's creation without being bound by their
-financial status. Just as how someone can go to the library and checkout any
-book that tickles their fancy, they should be freely able to learn from the
-ongoing research of our world. Of course one should recognize that not all
-information is safe or helpful to be freely available on the internet, but that
-is a topic for a later section.
+themselves by consuming quality research articles? All humans should be allowed to
+continue to learn more about God's creation without being bound by their financial status.
+The Bible encourages one to pursue knowledge. Proverbs 1:5 reads,
+\textquote[][.]{%
+  \textins{l}et the wise hear and increase in learning, and the one who
+  understands obtain guidance%
+}
+One's ability to live out this verse should not be artificially hindered by financial
+station. Just as how someone can go to the library and checkout any book that tickles
+their fancy, they should be freely able to learn from the ongoing research of our world.
+Of course one should recognize that not all information is safe or helpful to be freely
+available on the internet, but that is a topic for a later section.
 
@@
 cannot afford it, they might start to look for alternative way of obtaining it. But, when
-they can afford it and they think it is fairly priced, they will purchase it. Thus, piracy
-is often where tech-savvy persons turn to when they need something that they feel is
-exorbitantly priced, but is not often utilized when they feel the purchase is worthwhile.
-This way, piracy acts as a safeguard for consumers when they feel they are being taken
-advantage of by predatory business practices.
-% XXX the statements in the paragraph above are largely unsubstantiated >:D
-% XXX add darnton:pirating-and-publishing and bohannon:everyone sources if needed
+they can afford it and they think it is fairly priced, they will purchase it.
+
+According to a survey conducted by MUSO, \textcquote{muso:about}[,]{%
+  a data company that provides the most complete and trusted view of global
+  piracy and unlicensed media consumption%
+} the most common reasons that users pirate is \textcquote{vice:pirates}[.]{%
+  the content they were looking for wasn’t legally available (34 percent) was
+  too cumbersome or difficult to access (34 percent), or wasn’t affordable
+  (35 percent)%
+} The reason most people pirate is not that they are not willing to pay for the content,
+but that they cannot legally obtain it any other way, that it was too hard to access, or
+that it was not affordable. In the latter two cases, one can see the consumer's power to
+protest poorly designed systems or unaffordable prices. Thus, piracy is often where
+tech-savvy persons turn to when they need something that they feel is exorbitantly priced,
+but is not often utilized when they feel the purchase is worthwhile. This way, piracy
+acts as a safeguard for consumers when they feel they are being taken advantage of by
+predatory business practices.
 
@@
 for people in poor countries where piracy is widespread. Although they
-hypothesized, like other researchers, that \enquote{cyber criminals would target
+hypothesized, like other researchers, that \textquote{cyber criminals would target
 rich countries} because of monetary incentives, they found that, cyber criminals
@@
 Christians, we have a Lord that we can trust in with all our hearts (Prov. 3:5).
-As the Psalmist writes, \enquote{It is better to take refuge in the Lord than
+As the Psalmist writes, \textquote{It is better to take refuge in the Lord than
 to trust in man} (Ps. 118:8). Christians should be slow to put their entire

### 2024-04-29 paper.tex
Use \textins* vs \textins.
Some uses of \textins were technically supposed to be \textins*

@@
 \epigraph{%
-  \textins{T}hat ideas should freely spread from one to another over the globe,
+  \textins*{T}hat ideas should freely spread from one to another over the globe,
   for the moral and mutual instruction of man, and improvement of his
@@
 \textquote[][.]{%
-  \textins{l}et the wise hear and increase in learning, and the one who
+  \textins*{l}et the wise hear and increase in learning, and the one who
   understands obtain guidance%
@@
 Project writes, \textcquote{tor:about}[.]{%
-  \textins{t}he goal of onion routing was to have a way to use the internet with
+  \textins*{t}he goal of onion routing was to have a way to use the internet with
   as much privacy as possible, and the idea was to route traffic through
@@
 \textcquote[763]{thompson:trust}{%
-  \textins{s}uch blatant code would not go undetected for long. Even the most
+  \textins*{s}uch blatant code would not go undetected for long. Even the most
   casual perusal of the source of the C compiler would raise suspicions%
